---
title: "Climate Resilience: A Multi-Source Analysis of Climate Policy Evolution Post Paris Agreement (2013–2025)"
author: "Samantha Irving"
date: "`r Sys.Date()`"
abstract: "This application report documents a data mining pipeline which brings together four separate sources of climate data: the OECD CAPMF, the Climate Policy Database (CPDB), the UNFCCC NDCs and the Notre Dame Global Adaptation Initiative (ND-GAIN).The exploration analyzes the evolution of climate policy strategies since the signing of the Paris agreement on the 4th of November 2016, with a particular focus on the integration of adaptation measures alongside mitigation efforts. The focus is on a selection of 12 countries, with an effort to include members from both the Global South as well as the Global North. By combining API access, web scraping and natural language processing (NLP), the pipeline quantifies the shift from mitigation to adaptation and reveals a 'Commitment Gap' between Global North and Global South countries."
format: 
  pdf:
    toc: true
    toc-depth: 1
    
    number-sections: true
    colorlinks: true
    include-before-body:
      text: |
        \newpage
---

```{r "knitr config", include=FALSE}
# Chunk 1: JUST SET THE DIRECTORY
# This must be its own chunk with no other code
library(knitr)
# This finds your project root automatically
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```


```{r setup, include=FALSE}
# Load libraries and runmaster pipeline
# Make sure main.R is in the project root (one level up)


library(here)
source(here("main.R"))

```

# Introduction WIP

Climate policy research traditionally focuses on mitigation, with an emphasis on the reduction of greenhouse gas emissions. However, the increasing frequency and severity of climate impacts has elevated the importance of adaptation measures in national policy portfolios. The Paris Agreement (2015) marked a pivotal moment in global climate governance, with its dual focus on mitigation and adaptation, but measuring adaptation is inherently challenging. As the Grantham Research Institute notes, "unlike climate mitigation, where responses like carbon pricing and emission targets are widely tracked, adaptation laws and policies have received less systematic scrutiny" [1, p.10]. 

This measurement gap has real consequences. International climate databases systematically under represent adaptation policies, particularly in the Global South where adaptation is a **survival imperative** rather than a discretionary policy choice. When adaptation is invisible in the data, it becomes invisible in policy debates, climate finance allocation and accountability mechanisms. The lack of "comparative global analysis and systematic scrutiny of national adaptation legislation and policies has hampered the ability to assess progress, identify gaps, and understand effective frameworks" [1, p.10]. This project aims to bridge the measurement gap and serves as a digital audit of climate ambition across our selection of countries. 

Comparing the Global North to the Global South is a major focus of this project. The Global North has historically been responsible for the majority of emissions and has more resources to invest in both mitigation and adaptation. The Global South, whilst being less responsible for the acceleration of climate change, are the ones who bear the brunt of it. By analyzing the policy portfolios of countries from both groups, we can uncover disparities in climate ambition and identify areas where international support may be needed.

# Research Question WIP

By combining multiple data sources and methodologies, we seek to answer the question: *Have climate policy portfolios shifted from mitigation toward adaptation after the Paris Agreement, and does this shift differ between Global North and Global South countries?*

# Hypothesis WIP

# Data & Methodology WIP

I have structured the repository so that everything is managed by a single master script (`main.R`) that orchestrates the entire workflow, from data collection to analysis and visualization. This ensures that anyone can run the full pipeline with a single command and that all steps are transparent and reproducible.

The pipeline pulls from four distinct data environments:

- OECD CAPMF (API): Accessing standardized stringency indicators via the OECD R package.
- CPDB (Python Bridge): Utilizing reticulate to access the Climate Policy Database's unique policy-level data.
- UNFCCC NDCs (Web Scraping): An automated loop (documented in Issue #12) that identifies active English-language PDFs, downloads them to a local cache, and extracts raw text using pdftools.
- ND-GAIN (Manual/Hybrid): Due to dynamic web barriers on the Notre Dame site (Issue #19), this was handled as a manual download to ensure data integrity.



# Results WIP

## cpdb Cumulative Policy Adoption Over Time

The cumulative policy adoption plot (Figure 1) confirms the "Paris Effect." There is a visible acceleration in policy output for countries immediately following 2015.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-cumulative-stock
#| fig-cap: "Cumulative Climate Policy Adoption by Target Country (2012-2026)"
#| fig-width: 9
#| fig-height: 6
print(cpdb_cumulative_pol_adoption)
```

## The cpdb Adaptation Mix: Portfolio Composition Shifts
By calculating the Adaptation Mix, we see a striking North-South divide. Rwanda (RWA) and South Africa (ZAF) allocate a much higher percentage of their policy portfolio to adaptation than Germany (DEU) or Denmark (DNK).

Finding: Vulnerability (ND-GAIN) is a strong predictor of adaptation priority. As countries move into higher "Vulnerability Tiers," their legislative "Mindshare" shifts toward survival over emissions reduction.

Limitation: CPDB has much higher coverage for Global North countries, which may bias the analysis. The adaptation mix for some Global South countries may be underrepresented due to data gaps.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-adaptation-mix
#| fig-cap: "The Adaptation Mix: Percentage of Portfolio dedicated to Resilience"
print(cpdb_adaptation_mix_plot)
```


## The ndc Ambition Matrix: Lexical Hardness vs. Adaptation Focus

The Ambition Matrix is the main part of this project and consists of majority of the data work. By mapping Relative Adaptation Focus against Lexical Hardness (the ratio of "must/shall" vs "aim/should"), we see that:
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-ambition-matrix
#| fig-cap: "Climate Ambition Matrix: Lexical Hardness vs. Adaptation Focus"
print(ndc_ambition_vs_focus)
```


# Discussion WIP


# Limitations WIP


# Conclusion WIP

# Bibliography WIP

OECD (2025), The Climate Action Monitor 2025, OECD Publishing, Paris, https://doi.org/10.1787/1819c631-en.


# LLM Declaration WIP


**Word Count:** 

```{r}
#| echo: false
#| message: false
#| warning: false
# Calculate word count for the report
report_text <- readLines(here("04.report/Report.qmd"))
total_words <- sum(sapply(strsplit(report_text, "\\s+"), length))
cat("Total Word Count:", total_words)
```
